High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-19 notebooks/mnist.qmd
:sob

```{julia}
cov = .90
conf_model = conformal_model(mlp; method=:adaptive_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

@@

```{julia}
Random.seed!(1234)

# Set up search:
factual_label = 4
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 9
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 100

# Generate counterfactual using generic generator:
generator = GenericGenerator()
ce_wachter = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

# CCE:
λ=[0.0,1.0]
temp=0.5

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=λ, 
    temp=temp, 
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

@@

ces = zip([ce_wachter,ce_conformal,ce_conformal_jsma])
_names = ["Wachter", "CCE", "CCE-JSMA"]
counterfactuals = reduce((x,y)->cat(x,y,dims=3),map(ce -> CounterfactualExplanations.counterfactual(ce[1]), ces))
phat = reduce((x,y) -> cat(x,y,dims=3), map(ce -> target_probs(ce[1]), ces))
for x in zip(eachslice(counterfactuals; dims=3), _names, eachslice(phat; dims=3))
    ce, _name, _phat = (x[1],x[2],x[3])
    _title = "$_name (p̂=$(round(_phat[1]; digits=3)))"
    plt = Plots.plot(
        convert2image(MNIST, reshape(ce,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    plts = [plts..., plt]
end
plt = Plots.plot(plts...; size=(img_height*length(plts),img_height), layout=(1,length(plts)))
display(plt)
savefig(plt, joinpath(www_path, "cce_mnist.png"))
```

>>>>

```{julia}
cov = .90
conf_model = conformal_model(jem; method=:adaptive_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = CCE.ConformalModel(mach.model, mach.fitresult)
```

@@

```{julia}
# Random.seed!(1234)

# Set up search:
factual_label = 8
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 3
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 100

# Generate counterfactual using generic generator:
generator = GenericGenerator(opt=Flux.Optimise.Adam(),)
ce_wachter = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

generator = GreedyGenerator(η=1.0)
ce_jsma = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

# CCE:
λ=[0.0,1.0]
temp=0.01

# Generate counterfactual using CCE generator:
generator = CCEGenerator(
    λ=λ, 
    temp=temp, 
    opt=Flux.Optimise.Adam(),
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

@@

ces = [ce_wachter, ce_conformal, ce_jsma, ce_conformal_jsma]
_names = ["Wachter", "CCE", "JSMA", "CCE-JSMA"]
for x in zip(ces, _names)
    ce, _name = (x[1],x[2])
    x = CounterfactualExplanations.counterfactual(ce)
    _phat = target_probs(ce)
    _title = "$_name (p̂=$(round(_phat[1]; digits=3)))"
    plt = Plots.plot(
        convert2image(MNIST, reshape(x,28,28)),
        axis=nothing, 
        size=(img_height, img_height),
        title=_title
    )
    plts = [plts..., plt]
end
plt = Plots.plot(plts...; size=(img_height*length(plts),img_height), layout=(1,length(plts)))
display(plt)
savefig(plt, joinpath(www_path, "cce_mnist.png"))
```
